taxadb: A High-Performance Local Taxonomic Database Interface

### 2019-02-27 paper/manuscript.Rmd
exploring bird data in manuscript
and tweaking code along the way.

@@

```{r}
library(tidyverse)
library(taxadb)
```

```{r}
td_create("all")
```

```{r}
bbs <- read_tsv(system.file("extdata/bbs.tsv", package="taxadb"))
```

```{r}
dim(bbs)
```

```{r}
bbs_ids <- 
bbs %>% select(species) %>%
  mutate(gbif = get_ids(species, "gbif", "prefix"),
          col = get_ids(species, "col", "prefix"),
         itis = get_ids(species, "itis", "prefix"),
         ncbi = get_ids(species, "ncbi", "prefix"),
     wikidata = get_ids(species, "wd", "prefix"),
  #       iucn = get_ids(species, "iucn", "prefix"),
          ott = get_ids(species, "ott","prefix") 
  )

@@

```{r}
bbs_ott <- bbs$species %>% get_ids("ott", "prefix")
```

>>>>

```{r include=FALSE}
knitr::opts_chunk$set(message=FALSE)
```

@@

```{r message=FALSE}
library(tidyverse)
library(taxadb)
```

```{r message=FALSE}
td_create("all")
```

```{r message=FALSE}
bbs <- read_tsv(system.file("extdata/bbs.tsv", package="taxadb"))
```

```{r}
dim(bbs)

```{r}
bbs_ids <- 
bbs %>% 
  select(species) %>%
  mutate(
         gbif = get_ids(species, "gbif", "prefix"),
          col = get_ids(species, "col", "prefix"),
         itis = get_ids(species, "itis", "prefix"),
         ncbi = get_ids(species, "ncbi", "prefix"),
     wikidata = get_ids(species, "wd", "prefix"),
         iucn = get_ids(species, "iucn", "prefix"),
          ott = get_ids(species, "ott","prefix") 
  )

@@

```{r}
ott_ids <- bbs %>% 
  mutate(names = clean_names(species)) %>% 
  pull(names) %>% 
  ids("ott")

unmatched <- ott_ids %>% 
  filter(is.na(acceptedNameUsageID)) %>% 
  select(input, sort)
```

```{r}
unmatched$input %>% ids("iucn") %>% 
  select(scientificName, taxonomicStatus, input, taxonID, acceptedNameUsageID)

syn <- taxa_tbl("iucn") %>% 
  right_join(
    unmatched$input %>% 
    ids("iucn") %>% 
    select(acceptedNameUsageID),
    copy = TRUE) %>%
  select(scientificName, acceptedNameUsageID, taxonomicStatus) %>%
  collect()

## Some names have multiple synonyms, do not do this as a spread...    
syn %>% mutate(n = 1:length(scientificName)) %>% spread(taxonomicStatus, scientificName)
```

```{r}

## a name that resolves to multiple possible IDs is not considered a match by `get_ids()`
ott_ids <- bbs %>% 
  mutate(names = clean_names(species)) %>%
  pull(names) %>% 
  get_ids("ott", "prefix")

length(!is.na(ott_ids))
```
